---
title: "Arctic Cloud Dynamics"
authors: Aandishah Tehzeeb Samara, Serina Khalifa, Clare Gaffey
output: ioslides_presentation
---

```{r, eval = FALSE}
setwd(dir = "vignettes/slides")
rmarkdown::render(input = "Presentation.Rmd", output_file = "Presentation.html")
# This is to make the html output appear in the folder where you want it.
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Outline:
- Introduction (1-3 slides) (Serina, Clare (regression), Aandishah (editor))
- Data and Methods (2-4 slides) Methods (Aandhsish (1-2 + energy budget), 
- Results (2-4 slides) (Aandishah (2-3), Clare (1-2))
- Future work (1-3 slides) (Detrending?)


## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

## Introduction

## Data:
- NARR Reanalysis: (1979-2020)
  - Cloud cover at three altitude levels (low, medium, and high)
  - Wind speed at 10 m
  - Air temperature
  - Evaporation
  - Geopotential height
  - Relative humidity
- NASA MODIS-Aqua chlorophyll concentration (2002-2020)
- DoD SSMI/S sea ice concentration (1979-2020)


## Methods: Gradient Boosting Regression to find the drivers of low cloud cover concentration in the Chukchi Sea. {.smaller}
1. Monthly composites were downloaded (external/data/Download_Datasets.R)
2. Variables were converted to raster bricks, projected, cropped to the region
    of interest, and monthly averages were saved to a dataframe and exported
    to a csv to be combined into one master dataframe 
    (R/Prepare_Rasters_for_Model.R) [and Aandishah's combine code]
3. For running the gradient boosting model (XGBoost) the data were partitioned 
    into training and testing, trained a model, and then used accuracy 
    assessment to find the best number of trees for the final XGBoost model.

